---
title: "Workshop3real"
output: html_document
date: "2024-05-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)
```

```{r}
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```
#exercise
```{r}

```

```{r}
billboard |> 
  pivot_longer(
    cols = starts_with("wk"), 
    names_to = "week", 
    values_to = "rank"
  )
```

```{r}
billboard |> 
  pivot_longer(
    cols = starts_with("wk"), 
    names_to = "week", 
    values_to = "rank",
    values_drop_na = TRUE
  )
```

```{r}
df <- tribble(
  ~id,  ~bp1, ~bp2,
   "A",  100,  120,
   "B",  140,  115,
   "C",  120,  125
)
```


```{r}
df |> 
  pivot_longer(
    cols = bp1:bp2,
    names_to = "measurement",
    values_to = "value"
  )
```

```{r}
cms_patient_experience
```

```{r}
cms_patient_experience |> 
  distinct(measure_cd, measure_title)
```

```{r}
cms_patient_experience |> 
  pivot_wider(
    names_from = measure_cd,
    values_from = prf_rate
  )
```

```{r}
cms_patient_experience |> 
  pivot_wider(
    id_cols = starts_with("org"),
    names_from = measure_cd,
    values_from = prf_rate
  )
```

```{r}
df <- tribble(
  ~id, ~measurement, ~value,
  "A",        "bp1",    100,
  "B",        "bp1",    140,
  "B",        "bp2",    115, 
  "A",        "bp2",    120,
  "A",        "bp3",    105
)


```

```{r}
df |> 
  pivot_wider(
    names_from = measurement,
    values_from = value
  )
```

```{r}
df |> 
  distinct(measurement) |> 
  pull()
```

```{r}
df |> 
  select(-measurement, -value) |> 
  distinct()
```

```{r}
df |> 
  select(-measurement, -value) |> 
  distinct() |> 
  mutate(x = NA, y = NA, z = NA)
```

#exercise:
```{r}

stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")


```
#Why are pivot_longer() and pivot_wider() not perfectly symmetrical? 
#Because they serve different services and opparete in different ways
#They are designed to handle different types of data transformations. 

```{r}
table4a %>% 
  pivot_longer(c(1999, 2000), names_to = "year", values_to = "cases"))
```
#This code dosent work cause it has the wrong years and it misses one bracket at the end.


```{r}

preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)


```

#This table needs to be longer. This way, we can have a clearer view of the counts of males and females for each pregnancy status
#The three variables are pregnant, male, and female. 

```{r}
table3
```

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"))
```

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/")
```

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE)
```

```{r}
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```

```{r}
table5 %>% 
  unite(new, century, year, sep = "")
```
```{r}
treatment <- tribble(
  ~person,           ~treatment, ~response,
  "Derrick Whitmore", 1,         7,
  NA,                 2,         10,
  NA,                 3,         NA,
  "Katherine Burke",  1,         4
)
```

```{r}
treatment |>
  fill(everything())
```

```{r}
x <- c(1, 4, 5, 7, NA)
coalesce(x, 0)
```

```{r}
x <- c(1, 4, 5, 7, -99)
na_if(x, -99)
```

```{r}
x <- c(NA, NaN)
x * 10
#> [1]  NA NaN
x == 1
#> [1] NA NA
is.na(x)
#> [1] TRUE TRUE
```

```{r}
stocks <- tibble(
  year  = c(2020, 2020, 2020, 2020, 2021, 2021, 2021),
  qtr   = c(   1,    2,    3,    4,    2,    3,    4),
  price = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)
```

```{r}
stocks |>
  pivot_wider(
    names_from = qtr, 
    values_from = price
  )
```

```{r}
Student ID,Full Name,favourite.food,mealPlan,AGE 1,Sunil Huffmann,Strawberry yoghurt,Lunch only,4 2,Barclay Lynn,French fries,Lunch only,5 3,Jayendra Lyne,N/A,Breakfast and lunch,7 4,Leon Rossini,Anchovies,Lunch only, 5,Chidiegwu Dunkel,Pizza,Breakfast and lunch,five 6,Güvenç Attila,Ice cream,Lunch only,6
```

```{r}
? read_csv
students <- read_csv("C://data/students.csv")
```

```{r}
students <- read_csv("https://pos.it/r4ds-students-csv")
```

```{r}
students <- read_csv("data/students.csv", na = c("N/A", ""))

students
```

```{r}
students |> 
  rename(
    student_id = `Student ID`,
    full_name = `Full Name`
  )
```
#Exercise
```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")
```
#All the files are in Chr (character)

```{r}
library(tidyverse)
library(nycflights13)
```

```{r}
airlines
```

```{r}
airports
```

```{r}
planes
```

```{r}
weather
```

```{r}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)
```

```{r}
weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)
```

```{r}
flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)
```

```{r}
flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
```


```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
```

```{r}
flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
```


```{r}
flights2 %>%
  select(-origin, -dest) %>% 
  mutate(name = airlines$name[match(carrier, airlines$carrier)])
```

```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     3, "x3"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     4, "y3"
)
```

```{r}
x %>% 
  inner_join(y, by = "key")
```

```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     1, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2"
)
left_join(x, y, by = "key")
```

```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     3, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     2, "y3",
     3, "y4"
)
left_join(x, y, by = "key")
```
```{r}
flights2 %>% 
  left_join(weather)
```

```{r}
flights2 %>% 
  left_join(planes, by = "tailnum")
```

```{r}
flights2 %>% 
  left_join(airports, c("dest" = "faa"))
```

```{r}

flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```



